---
title: "SBL_analysis"
author: "Steven Lawrence"
date: "September 13, 2019"
output: github_document
---

#Data Retrival and Manipulation

Here I am installing all libraries neccessary for analysis

```{r Libraries Install, echo=TRUE, eval=FALSE, message=FALSE}
install.packages(c("tidyverse","ggplot2","tseries","DataCombine","forecast","dplyr","knitr","pastecs","grid","formattable","gridExtra","viridis","viewpoint"))

```
Loading Libraries

```{r Libraries, message=FALSE, warning=FALSE}
library(tidyverse)
library(tseries)
library(DataCombine)
library(forecast)
library(knitr)
library(pastecs)
library(grid)
library(formattable)
library(gridExtra)
library(viridis)
library(magrittr)
library(qwraps2)

```


##Geographic Analysis 


##Temporal Trend
Both temporal datasets were retrieve from trends.google.com via csv files and are depicted below showing the head and tail of each. 

```{r Skin Bleaching Temoral Datasets, message=FALSE, echo=FALSE, warning=FALSE}
skBT<- read.csv("C:/Users/brian/Documents/Biostatistis Mount Sinai/SBL-Global/drive-download-20190913T154116Z-001/GtrendData/SkinBleachingTime0419.csv", header = T, sep = ",",stringsAsFactors = F, col.names = c("Time","RSV"))
skBT<- skBT[-1,]
skBT$RSV<- as.numeric(skBT$RSV)
```

Head and Tail of Skin Bleaching temporal trend dataset
```{r SKBT head and tail ,message=FALSE, warning=FALSE, echo=FALSE}
head(skBT)
tail(skBT)
```

```{r Skin Lightening Temoral Datasets, message=FALSE, echo=FALSE, warning=FALSE}
skLT<- read.csv("C:/Users/brian/Documents/Biostatistis Mount Sinai/SBL-Global/drive-download-20190913T154116Z-001/GtrendData/SkinLighteningTime0419.csv", header = T, sep = ",", stringsAsFactors = F, col.names = c("Time","RSV"))
skLT<- skLT[-1,]
skLT$RSV<- as.numeric(skLT$RSV)
```
Head and tail of skin lightening temporal trend dataset
```{r SKLt head and tail ,message=FALSE, warning=FALSE, echo=FALSE}

head(skLT)
tail(skLT)
```
Here I merged the two temporal datasets to create a new variable that is the average of the two.  The head and tail is shown below.

```{r Merged Temporal Trend Data, message=FALSE, warning=FALSE}

## merging bleaching and lightening data temporal data sets relative search volumes by time##

SKBLT<- merge(skBT, skLT, by = "Time", sort = T)

colnames(SKBLT)<- c("Date","Bleaching", "Lightening")

#creating an average variable of bleaching and lightening temporal RSV's

SKBLT$aveRSV<- (SKBLT$Bleaching+SKBLT$Lightening)/2
#creating a subset of SKBLT only time and average RSV 
SKBLT.p<- SKBLT[,c(1,4)]
head(SKBLT.p)
tail(SKBLT.p)
```

Using the tseries and forecast package we analyazed the temporal trend. 

```{r time series object, message=FALSE, warning=FALSE}
#creating a time series object in order to plot the temporal trend as a time series element
SKBLT.ts<- ts(SKBLT.p$aveRSV)

```

This plot is SBL queries by month in the United States from 2004 - 2019. 
```{r temporal trend plot, message=FALSE, warning=FALSE}
#plotting sbl quries over time with a resolution of months 
plot(SKBLT.ts , type = "o", main = "SKin Bleaching and Lightening Trend from 2004 To 2019", xlab = "Months", ylab = "Relative Search Volume")
```
Here I provide some discriptive statistics of the overall temporal trend. 

Then data is not normally distributed there for I recorded the median and IQR `median_iqr(SKBLT.p$aveRSV)`
```{r Discriptibe Statisics}
hist(SKBLT.p$aveRSV, n = 40)
mean_skbl<-mean(SKBLT.p$aveRSV)
mean_skbl


```
SKBLT.pc<-PercChange(SKBLT.p, Var = "aveRSV", slideBy = -1)
SKBLT.pc
summary(SKBLT.pc$`aveRSV_PChangeFrom-1`)
#   Min. 1st Qu.  Median    Mean 3rd Qu.    Max.    NA's 
#-49.689  -9.184   1.449   2.661   9.630 111.268       1 

SKBLT.p9<- SKBLT.p[-c(1:60),]
SKBLT.pc9<-PercChange(SKBLT.p9, Var = "aveRSV", slideBy = -1)
SKBLT.pc9[is.na(SKBLT.pc9)]<- 0 
summary(SKBLT.pc9$`aveRSV_PChangeFrom-1`)
#     Min.  1st Qu.   Median     Mean  3rd Qu.     Max. 
#-49.6894  -7.1545   0.8283   1.9474   8.7086 111.2676
#########################

#Checking seasonality
count_ma<- ts(na.omit(SKBLT.p$aveRSF), frequency = 12)
decomp<-stl(count_ma, s.window = "periodic")
plot(decomp)
decomp<- as.ts(decomp)
decomp
deseason<- seasadj(decomp)

install.packages("forecast")
library(forecast)
autoplot(decomp)
ggseasonplot(deseason, year.lables = T, continuous = T, polar = T)

#########################

Creating a data frame of the non white population which will be inverted values of the white population scores. 
```{r}
Region<- c("Alabama", "Alaska","Arizona", "Arkansas", "California", "Colorado","Connecticut", "Delaware", "District of Columbia", "Florida", "Georgia", "Hawaii", "Idaho", "Illinois", "Indiana", "Iowa", "Kansas", "Kentucky", "Louisiana", "Maine", "Maryland", "Massachusetts", "Michigan", " Minnesota", "Mississippi", "Missouri", "Montana", "Nebraska", "Nevada","New Hampshire", "New Jersey", "New Mexico", "New York", "North Carolina", "North Dakota", "Ohio", "Oklahoma", "Oregon", "Pennsylvania", "Puerto Rico", "Rhode Island", "South Carolina", " South Dakota", "Tennessee", "Texas", "Utah", "Vermont", " Virginia", "Washington", "West Virginia", "Wisconsin", "Wyoming")

wonly<- c(65.4,60.3,54.4,72.2,36.8,67.9,66.5,61.9,37.1,53.5,52.4,21.8,81.7,61.0,78.9,85.3,75.7,84.3, 58.6,93.1,50.5,71.4,74.9,79.5,56.5, 79.3,85.9,78.6,48.7,90.0,54.9,37.1,55.4,62.8,84.0,78.7,65.3,75.3,76.1,0.8,72.0,63.7,81.4,73.7,41.5,78,92.5,61.5,68.0,92.1,81.1,83.8 )

woPop<- data.frame(Region,wonly)
woPop %>% mutate_if(is.factor, as.character) -> woPop


```
skblr.naomit

skblmmp<- merge(woPop, skblr.naomit, by ="Region", all = T )
skblmmp

skblmmp.naomit<- na.omit(skblmmp)

skblmmp.naomit<- skblmmp.naomit[order(-skblmmp.naomit$aveRSF),]

skblmmp.naomit$nwp<- 100 - skblmmp.naomit$wonly

skblmmp.naomit<- skblmmp.naomit[order(-skblmmp.naomit$nwp),]
skblmmp.naomit

skblm1<- as_tibble(skblmmp.naomit)
skblm1<- skblm1%>%dplyr::mutate(region=stringr::str_to_lower(Region))
sMap<- ggplot2::map_data("state")
#skblmm = skin bleaching lightening merge map
skblmm1<- merge(sMap, skblm1, by = 'region')
skblmm1<- skblmm1%>% dplyr::left_join(x=., y = sMap, by ='region')
skblmm1
skblmm1<- skblmm1[,-c(7,)]
colnamesskblmm1

skblmm1%>% ggplot(aes(x=long, y = lat.y))+ geom_polygon(aes(group = group.y, fill = log(nwp)), colour = 'white')+theme2()+ggtitle("Skin Bleaching and Skin Lightening by Region")


Extracting Census data
